Analyzing Information Transfer from Heterogeneous Sources via Precise High-dimensional Asymptotics

### 2021-07-25 revision_journal/abstract.tex
abstract

\begin{abstract}
%	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

>>>>

\begin{abstract}
    We consider the problem of transfer learning---gaining knowledge from one task and applying it to a different but related target task.
    A fundamental question in transfer learning is when combining the data of both tasks works better than using only the target task's data (equivalently, when there is a ``positive information transfer'').
    We study this question formally in a linear regression setting where a certain two-layer linear neural network estimator is used to combine both tasks' data.
    Under various settings involving covariate and model shifts, the estimator reproduces several interesting behaviors regarding positive (and negative) information transfer observed on real-world data.

    We then show the precise asymptotic limit of risk of the estimator in terms of geometries of the data in a high-dimensional setting, where the sample sizes increase with the feature dimension proportionally at a fixed ratio.
    We also find that the asymptotics is accurate for finite dimensions and apply them to analyze the exact conditions for positive information transfer.
    This analysis leads to several novel insights regarding covariate and model shifts.
    For example, the risk curve is non-monotone under model shift, thus motivating a more efficient procedure for progressively adding data.
    The main ingredient of our analysis is showing the asymptotic limit of certain combinations of independent sample covariance matrices under covariate shift and different sample sizes, which may be of independent interest.

%	We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio.  Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

### 2021-07-25 revision_journal/exp.tex
finish intro
abstract

\begin{figure}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_covariate_shift.eps}
		\caption{Covariate shift}
		\label{fig_sec5_covariate}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_model_shift.eps}
		\caption{Model shift}
		\label{fig_sec5_model}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_cov_model_shift.eps}
		\caption{Covariate and model shift}
		\label{fig_sec5_cov_model}
	\end{subfigure}
	\caption{We compare the empirical excess risk of HPS and several natural transfer learning estimators. We find that HPS achieves the lowest excess risk for predicting task two under (either or both) covariate and model shift. This simulation uses $p = 50, n_1 = n_2 = 100, \sigma = 1/2$. Figure \ref{fig_sec5_cov_model} uses $\lambda = 4$ for generating covariate-shifted features.}
	\label{fig_sec51}
\end{figure}

@@

\begin{wrapfigure}[12]{r}{0.3\textwidth}
	\vspace{-0.35in}
	\begin{center}
		\includegraphics[width=0.27\textwidth]{figures/mtl_model_arch.pdf}
	\end{center}
	\vspace{-0.28in}
	\caption{A hard parameter sharing neural network.}
	\label{fig_intro_arch}
\end{wrapfigure}
We conduct further studies of HPS for text classification tasks.
We consider six datasets for predicting movie review sentiment (MR and SST), sentence subjectivity (SUBJ), customer review sentiment (CR), question types (TREC), and polarity of a phrase (MPQA).\footnote{The datasets can be downloaded at \url{https://github.com/harvardnlp/sent-conv-torch/tree/master/data}. Further statistics of each dataset can also be found following the link.}
%\cite{pang2005seeing,socher2013recursive}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
% \cite{lei2018simple}
%The question is to predict positive or negative sentiment expressed in the text.
We learn a hard parameter sharing model that consists of a word embedding layer using GloVe \cite{pennington2014glove}, followed by a shared feature representation layer.\footnote{The GloVe word vectors can be downloaded at \url{https://nlp.stanford.edu/projects/glove/}.}
A separate output layer is used for predicting the labels of each dataset.
We evaluate on three possibilities of feature representation layers including LSTM, MLP, and CNN, as implemented by PyTorch.\footnote{\url{pytorch} Before the MLP layer, we apply an average pooling layer over the embedding of every word in the sentence.}
See Figure \ref{fig_intro_arch} for an illustration of the network architecture.

>>>>

\begin{figure}[!t]
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_covariate_shift.eps}
		\caption{Covariate shift}
		\label{fig_sec5_covariate}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_model_shift.eps}
		\caption{Model shift}
		\label{fig_sec5_model}
	\end{subfigure}\hfill%
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.95\textwidth]{figures/compare_risk_cov_model_shift.eps}
		\caption{Covariate and model shifts}
		\label{fig_sec5_cov_model}
	\end{subfigure}
	\caption{We compare the empirical excess risk of HPS and several natural transfer learning estimators. We find that HPS achieves the lowest excess risk for predicting task two under covariate and model shifts. This simulation uses $p = 50, n_1 = n_2 = 100, \sigma = 1/2$. Figure \ref{fig_sec5_cov_model} uses $\lambda = 4$ for generating covariate-shifted features.}
	\label{fig_sec51}
\end{figure}

@@

\begin{wrapfigure}[11]{r}{0.3\textwidth}
	\vspace{-0.15in}
	\begin{center}
		\includegraphics[width=0.25\textwidth]{figures/mtl_model_arch.pdf}
	\end{center}
	\vspace{-0.28in}
	\caption{A hard parameter sharing neural network.}
	\label{fig_intro_arch}
\end{wrapfigure}
We conduct further studies of HPS for text classification tasks.
We consider six datasets for predicting movie review sentiment (MR and SST), sentence subjectivity (SUBJ), customer review sentiment (CR), question types (TREC), and polarity of a phrase (MPQA).\footnote{The datasets can be downloaded at \url{https://github.com/harvardnlp/sent-conv-torch/tree/master/data}. Further statistics of each dataset can also be found following the link.}
%\cite{pang2005seeing,socher2013recursive}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
%\cite{pang2004sentimental}
% \cite{lei2018simple}
%The question is to predict positive or negative sentiment expressed in the text.
We learn a hard parameter sharing model that consists of a word embedding layer using GloVe \cite{pennington2014glove}, followed by a shared feature representation layer.\footnote{The GloVe word vectors can be downloaded at \url{https://nlp.stanford.edu/projects/glove/}.}
A separate output layer is used for predicting the labels of each dataset.
We evaluate on three possibilities of feature representation layers including LSTM, MLP, and CNN, all of which are implemented using PyTorch.\footnote{The PyTorch neural network modules are available at \url{https://pytorch.org/docs/stable/nn.html}.}
See Figure \ref{fig_intro_arch} for an illustration of the network architecture.

### 2021-07-25 revision_journal/intro.tex
abstract

Given data from two related tasks, how much does combining both data for learning a joint model help predict one of the tasks of interest?
For example, suppose you have $n_1$ datapoints with $p$-dimensional covariates (sampled from a distribution $D_1$) and real-valued labels in the first task, and $n_2$ datapoints with $p$-dimensional covariates (sampled from a different distribution $D_2$) and real-valued labels in the second task. 
Does combining the $n_1 + n_2$ datapoints help learn a model whose performance is better than a model learned using the $n_2$ datapoints alone?
Such questions are prevalent in transfer learning, where one often uses data from related tasks to expand the sample size of a task of interest.
A frequent impediment to applying transfer learning is \textit{negative information transfer} if combining both data performs worse than learning a single task clone \cite{PY09}.
\textit{Positive information transfer} refers to the desirable scenario where a task of interest benefits when combined with another task.

@@

The key ingredient of our analysis is showing the asymptotic limit of certain combinations of sample covariance matrices under distribution shift, using the local laws developed recently by \citet{isotropic} and \citet{BES_free1}.
After formally defining the specific problem that we tackle, in Section \ref{sec_related}, we provide a technical discussion of the related works in random matrix theory.

@@

Let $i = 1$ or $2$.
Let $x_1^{(i)}, x_2^{\ss (i)}, \dots, x_{n_i}^{(i)}$ be the covariates.
Let $y_1^{(i)}, y_2^{(i)}, \dots, y_{n_i}^{(i)}$ be the corresponding labels.
We assume a linear model specified by an unknown model vector $\beta^{(i)} \in \real^p$ as follows:
\begin{align}
    y_j^{(i)} = \inner{x_j^{(i)}}{\beta^{(i)}} + \varepsilon_j^{(i)}, \text{ for any } j = 1,\dots,n_i, \label{eq_linear}
\end{align}
where $\varepsilon^{(i)}_j\in\real$ denotes a random noise variable with mean zero and variance $\sigma^2$.
For the ease of presentation, we refer to the first dataset as the source and the second task as the target.
%Our goal is to learn an estimator to predict the target task given both datasets.

We learn a two-layer linear neural network with parameters $A \in \real^{p\times r}$, $B_1 \in \real^{r}$, and $B_2 \in \real^r$ by minimizing the following optimization objective:
\begin{align}\label{eq_hps}
    f(A, B_1, B_2) = \bignorm{X^{(1)} A B_1 - Y^{(1)}}_2^2 + \bignorm{X^{(2)} A B_2 - Y^{(2)}}_2^2.
\end{align}
The above objective uses $A$ as the shared feature space for both tasks and $B_1, B_2$ as the prediction head for each task.\footnote{Popular variants of the objective function \eqref{eq_hps} include adding a weight parameter to each summand and adding ridge regularization. We focus on the unweighted and unregularized objective for simplicity. All of the asymptotic limits we show below can be straightforwardly extended to weighted and regularized objectives.}
Such models are also known as \textit{hard parameter sharing} \cite{C97,R17}.
We focus on the case of $r = 1$---otherwise, the global minimizer of $f(A, B_1, B_2)$ reduces to learning each task alone, resulting in no transfer effect (cf. Proposition \ref{prop_large_r} in Section \ref{sec_risk}).
In this case, let $(\hat{A}, \hat{B}_1, \hat{B}_2)$ be a global minimizer of $f(\cdot)$.
The hard parameter sharing (HPS) estimator for the target task is defined as $\hat{\beta}_2^{\MTL} \define \hat{B} \hat{A}_2$.
In order to evaluate the estimator, we will study the excess risk, which is the prediction risk at an unseen datapoint (of task two) minus $\sigma^2$, denoted as $L(\hat{\beta}_2^{\MTL})$.

@@

We consider a high-dimensional setting on each dataset where the sample size $n_i$ increases with $p$ to infinity proportionally by a constant factor, which has recently received lots of interests (see e.g., \citet{hastie2019surprises} and Section \ref{sec_related} for more references).
Let $\Sigma^{(i)} \in \real^{p\times p}$ be a deterministic positive semidefinite matrix.
The covariates are assumed to have population covariance equal to $\Sigma^{(i)}$: $x_j^{(i)} = \big(\Sigma^{(i)}\big)^{1/2} z_{j}^{(i)},$ for any $j = 1,\dots,n_i,$
where $z_j^{(i)}$ consists of independent and identical entries sampled from a distribution that has zero mean, unit variance, and satisfies certain bounded moment condition (see Section \ref{sec_data} for the precise assumptions).

\paragraph{Main results.} We focus on the so called \textit{underparametrized} setting for the target task where $n_2 > p$ whereas for dataset one $n_1$ can be either smaller or greater than $p$.
Thus, the OLS estimator exists and it is well-known that the limit of its excess risk when $p$ approaches infinity is equal to $\frac{\sigma^2 p}{n_2 - p}$ (see e.g., \citet{bai2009spectral}).
By contrast, showing the asymptotic limit of $L(\hat{\beta}^{\MTL}_2)$ in the high-dimensional setting requires dealing with the inverse of the sum of two sample covariance matrices under distribution shift.
%The first challenge is dealing with \textit{covariate shift}---when $\Sigma^{(1)}$ and $\Sigma^{(2)}$ are different.
%The second challenge is dealing with \textit{model shift}---when $\beta^{(1)}$ and $\beta^{(2)}$ are different.
Our main result addresses both challenges by \todo{illustrate}.
\begin{itemize}
    \item %\textbf{Covariate shift:}
    Our first result applies to settings where $\Sigma^{(1)}$ and $\Sigma^{(2)}$ can be arbitrarily different but $\beta^{(1)} = \beta^{(2)}$.
    In Theorem \ref{thm_main_RMT}, we characterize the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ as a function of the singular values of the matrix $\big(\Sigma^{(1)}\big)^{1/2}\big(\Sigma^{(2)}\big)^{-1/2}$ and the sample sizes $n_1, n_2$.
    \item %\textbf{Model shift:}
    Our second result applies to settings where $\beta^{(1)}$ and $\beta^{(2)}$ can be arbitrarily different but $\Sigma^{(1)} = \Sigma^{(2)}$.
    In Theorem \ref{cor_MTL_loss}, we provide the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ as a function of the $\ell_2$ distance between $\beta^{(1)}$ and $\beta^{(2)}$, and the sample sizes $n_1, n_2$.
    \item %\textbf{Covariate and model shift:}
    Our third result applies to settings where both the covariance matrices and the model vectors are different between the two tasks.
    In Theorem \ref{???}, we show that when $\Sigma^{(2)}$ is isotropic, the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ is characterized by the singular values of $\Sigma^{(1)}$, the $\ell_2$ distance between $\beta^{(1)}$ and $\beta^{(2)}$), and the sample sizes $n_1, n_2$.
    Next, in Theorem \ref{prop_main_RMT}, we consider the most general setting concerning an arbitrary $\Sigma^{(2)}$. We identify an estimate of $L(\hat{\beta}_2^{\MTL})$ that becomes more and more accurate as $n_1 / n_2$ increases.
\end{itemize}

@@

\paragraph{Theoretical insights and empirical implications.} Next, we demonstrate the power of these precise asymptotics by explaining several empirical phenomena in transfer learning.
We consider a random-effect model where $\beta^{(i)}$ is equal to a shared model vector plus per-task variances (cf. \citet{dobriban2018high}).
Under this model, we show the following \textit{theoretical insights}.
\begin{enumerate}
    \item[i)] \textit{Whether or not covariate shift helps depends on sample sizes:} We consider the covariate shift setting and show that (cf. Proposition \ref{claim_dichotomy}) when $n_1 < n_2$, transferring from a data source with $\Sigma^{(1)} \neq \Sigma^{(2)}$ can achieve a smaller $L(\hat{\beta}_2^{\MTL})$ compared to transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$. When $n_1 > n_2$, transferring from from a data source with $\Sigma^{(1)} \neq \Sigma^{(2)}$ always results in higher $L(\hat{\beta}_2^{\MTL})$ than transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$.
    \item[ii)] \textit{Identical covariance provides approximately optimal transfer under imbalanced sizes:} Additionally, in Proposition \ref{prop_covariate}, we show that when $n_1$ is greater than $n_2$ times a certain large constant factor, transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$ achieves an approximately minimum excess risk value, within a certain family of covariance matrices of $\Sigma^{(1)}$.
    \item[iii)] \textit{Three regimes of information transfer under model shift}: Next, we consider the model shift setting and characterize the precise range of $n_1$ under which there is a positive transfer.
    In Proposition \ref{claim_model_shift}, we show that when $\mu^2 \le \frac{\sigma^2 p}{2(n_2 - p)}$ (recall that $\norm{\beta^{(1)} - \beta^{(2)}}^2 \approx 2\mu^2$) is small, the transfer effect is positive for any $n_1$.
    When $\frac{\sigma^2 p}{2(n_2 - p)} < \mu^2 < \frac{\sigma^2 n_2}{2(n_2 - p)}$, there is a transition from positive to negative transfer, as $n_1$ increases.
    When $\frac{\sigma^2 n_2}{2(n_2 - p)} \le \mu^2$, the transfer effect is negative for any $n_1$.
    Additionally, we illustrate that these three regimes of information transfer arises under both covariate and model shift (cf. Figure \ref{fig_sec3_cov_mo_b}).
\end{enumerate}

>>>>

Given data from two related tasks, how much does combining both data for learning a joint model help predict one of the tasks of interest?
For example, suppose you have $n_1$ datapoints with $p$-dimensional covariates (sampled from a distribution $D_1$) and real-valued labels in the first task, and $n_2$ datapoints with $p$-dimensional covariates (sampled from a different distribution $D_2$) and real-valued labels in the second task. 
Does combining the $n_1 + n_2$ datapoints help learn a model whose performance is better than a model learned using the $n_2$ datapoints alone?
Such questions are prevalent in transfer learning, where one often uses data from related tasks to expand the sample size of a task of interest.
A frequent impediment to applying transfer learning is \textit{negative information transfer} if combining both data performs worse than learning a single task clone \cite{PY09}.
\textit{Positive information transfer} refers to the desirable scenario where a task of interest benefits from another task's data.

@@

The key ingredient of our analysis is showing the asymptotic limit of certain combinations of sample covariance matrices under distribution shift, using the \textit{local laws} developed recently by \citet{isotropic} and \citet{BES_free1}.
After formally defining the specific problem that we tackle, in Section \ref{sec_related}, we provide a technical discussion of the related works in random matrix theory.

@@

Let $i = 1$ or $2$.
Let $x_1^{(i)}, x_2^{\ss (i)}, \dots, x_{n_i}^{(i)}$ be the covariates.
Let $y_1^{(i)}, y_2^{(i)}, \dots, y_{n_i}^{(i)}$ be the corresponding labels.
We assume a linear model specified by an unknown model vector $\beta^{(i)} \in \real^p$ as follows:
\begin{align}
    y_j^{(i)} = \inner{x_j^{(i)}}{\beta^{(i)}} + \varepsilon_j^{(i)}, \text{ for any } j = 1,\dots,n_i, \label{eq_linear}
\end{align}
where $\varepsilon^{(i)}_j\in\real$ denotes a random noise variable with mean zero and variance $\sigma^2$.
We focus on the so called \textit{underparametrized} setting for the target task where $n_2 > p$ whereas for dataset one $n_1$ can be either smaller or greater than $p$  (see Section \ref{sec_data} for the precise assumptions).
For the ease of presentation, we refer to the first task as the source and the second task as the target.

We learn a two-layer linear neural network with parameters $A \in \real^{p\times r}$, $B_1 \in \real^{r}$, and $B_2 \in \real^r$ by minimizing the following optimization objective:
\begin{align}\label{eq_hps}
    f(A, B_1, B_2) = \bignorm{X^{(1)} A B_1 - Y^{(1)}}_2^2 + \bignorm{X^{(2)} A B_2 - Y^{(2)}}_2^2.
\end{align}
The above objective uses $A$ as the shared feature space for both tasks and $B_1, B_2$ as the prediction head for each task.\footnote{Popular variants of the objective function \eqref{eq_hps} also involve weight parameters for each task and ridge regularization. We focus on the unweighted and unregularized objective for simplicity. All of the asymptotic limits described below can be straightforwardly extended to weighted and regularized objectives.}
Such models are also known as \textit{hard parameter sharing} \cite{C97,R17}.
We focus on the case of $r = 1$---otherwise, the global minimizer of $f(A, B_1, B_2)$ reduces to learning each task alone, resulting in zero information transfer (cf. Proposition \ref{prop_large_r} in Section \ref{sec_risk}).
For the case of $r = 1$, let $(\hat{A}, \hat{B}_1, \hat{B}_2)$ be $f(\cdot)$'s global minimizer, which is unique because $n_2 > p$.
The hard parameter sharing (HPS) estimator for the target task is defined as $\hat{\beta}_2^{\MTL} \define \hat{B} \hat{A}_2$.
In order to evaluate the estimator, we will study the excess risk, which is the prediction risk at an unseen datapoint (of task two) minus $\sigma^2$, denoted as $L(\hat{\beta}_2^{\MTL})$.

@@

We consider a high-dimensional setting on each dataset where the sample size $n_i$ increases with $p$ to infinity proportionally by a constant factor, which has recently received lots of interests (see e.g., \citet{hastie2019surprises} and Section \ref{sec_related} for more references).
Let $\Sigma^{(i)} \in \real^{p\times p}$ be a deterministic positive semidefinite matrix.
The covariates are assumed to have population covariance equal to $\Sigma^{(i)}$: $x_j^{(i)} = \big(\Sigma^{(i)}\big)^{1/2} z_{j}^{(i)},$ for any $j = 1,\dots,n_i,$
where $z_j^{(i)}$ consists of independent and identical entries sampled from a distribution that has zero mean, unit variance, and satisfies certain bounded moment condition.

\paragraph{Main results.} 
Thus, the OLS estimator exists and it is well-known that the limit of its excess risk when $p$ approaches infinity is equal to $\frac{\sigma^2 p}{n_2 - p}$ (see e.g., \citet{bai2009spectral}).
By contrast, showing the asymptotic limit of $L(\hat{\beta}^{\MTL}_2)$ in the high-dimensional setting requires dealing with the inverse of the sum of two sample covariance matrices under distribution shift.
%The first challenge is dealing with \textit{covariate shift}---when $\Sigma^{(1)}$ and $\Sigma^{(2)}$ are different.
%The second challenge is dealing with \textit{model shift}---when $\beta^{(1)}$ and $\beta^{(2)}$ are different.
Our main result addresses both challenges by \todo{illustrate}.
\begin{itemize}
    \item %\textbf{Covariate shift:}
    Our first result applies to settings where $\Sigma^{(1)}$ and $\Sigma^{(2)}$ can be arbitrarily different but $\beta^{(1)} = \beta^{(2)}$.
    In Theorem \ref{thm_main_RMT}, we characterize the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ as a function of the singular values of the matrix $\big(\Sigma^{(1)}\big)^{1/2}\big(\Sigma^{(2)}\big)^{-1/2}$ and the sample sizes $n_1, n_2$.
    \item %\textbf{Model shift:}
    Our second result applies to settings where $\beta^{(1)}$ and $\beta^{(2)}$ can be arbitrarily different but $\Sigma^{(1)} = \Sigma^{(2)}$.
    In Theorem \ref{cor_MTL_loss}, we provide the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ as a function of the $\ell_2$ distance between $\beta^{(1)}$ and $\beta^{(2)}$, and the sample sizes $n_1, n_2$.
    \item %\textbf{Covariate and model shift:}
    Our third result applies to settings where both the covariance matrices and the model vectors are different between the two tasks.
    In Theorem \ref{???}, we show that when $\Sigma^{(2)}$ is isotropic, the asymptotic limit of $L(\hat{\beta}_2^{\MTL})$ is characterized by the singular values of $\Sigma^{(1)}$, the $\ell_2$ distance between $\beta^{(1)}$ and $\beta^{(2)}$), and the sample sizes $n_1, n_2$.
    Next, in Theorem \ref{prop_main_RMT}, we consider the most general setting concerning an arbitrary $\Sigma^{(2)}$. We identify an estimate of $L(\hat{\beta}_2^{\MTL})$ that becomes more and more accurate as $n_1 / n_2$ increases.
\end{itemize}

@@

\paragraph{Theoretical insights and empirical implications.} Next, we demonstrate the power of these precise asymptotics by explaining several empirical phenomena in transfer learning.
We consider a random-effect model where $\beta^{(i)}$ is equal to a shared model vector plus per-task variances (cf. \citet{dobriban2018high}).
Under this model, we show the following \textit{theoretical insights}.
\begin{enumerate}
    \item[i)] \textit{Whether or not covariate shift helps depends on sample sizes:} We consider the covariate shift setting and show that (cf. Proposition \ref{claim_dichotomy}) when $n_1 < n_2$, transferring from a data source with $\Sigma^{(1)} \neq \Sigma^{(2)}$ can achieve a smaller $L(\hat{\beta}_2^{\MTL})$ compared to transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$. When $n_1 > n_2$, transferring from from a data source with $\Sigma^{(1)} \neq \Sigma^{(2)}$ always results in higher $L(\hat{\beta}_2^{\MTL})$ than transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$.
    \item[ii)] \textit{Identical covariance provides approximately optimal transfer under imbalanced sizes:} Additionally, in Proposition \ref{prop_covariate}, we show that when $n_1$ is greater than $n_2$ times a certain large constant factor, transferring from a data source with $\Sigma^{(1)} = \Sigma^{(2)}$ achieves an approximately minimum excess risk value, within a certain family of covariance matrices of $\Sigma^{(1)}$.
    \item[iii)] \textit{Three regimes of information transfer under model shift}: Next, we consider the model shift setting and characterize the precise range of $n_1$ under which there is a positive transfer.
    In Proposition \ref{claim_model_shift}, we show that when $\mu^2 \le \frac{\sigma^2 p}{2(n_2 - p)}$ (recall that $\norm{\beta^{(1)} - \beta^{(2)}}^2 \approx 2\mu^2$) is small, the transfer effect is positive for any $n_1$.
    When $\frac{\sigma^2 p}{2(n_2 - p)} < \mu^2 < \frac{\sigma^2 n_2}{2(n_2 - p)}$, there is a transition from positive to negative transfer, as $n_1$ increases.
    When $\frac{\sigma^2 n_2}{2(n_2 - p)} \le \mu^2$, the transfer effect is negative for any $n_1$.
    Figure \ref{fig_motivation} suggests that the asymptotic limits are also accurate in finite dimensions.
    Additionally, these three regimes of information transfer can be observed in settings of covariate and model shifts (cf. Figure \ref{fig_sec3_cov_mo_b}).
\end{enumerate}
